DATA DRIVEN MODELS FOR APOGEE M DWARFS

### 2018-03-27 paper/draft1/draft1.tex
update data paragraph

\usepackage{color}
\usepackage{amsmath}
\usepackage{enumitem}

@@

The APOGEE survey \citep{Majewski:2015} of the Sloan Digital Sky Survey IV (SDSS-IV) is a high resolution (R$\sim$22,500), H-band (1.5-1.7$\mu$m) survey which has observed over 260,000 stellar spectra since its fourteenth data release (DR14; \citealt{Abolfathi:2017}). While APOGEE is primarily targeted for bright stellar populations (descirbe cuts \citealt{Zasowski:2013}) with the objective of studying galactic composition and evolution, numerous cool, main sequence have also been observed either as targets proposed by the M dwarf ancilliary survey \citep{Desphande:2013}

Fundamental parameters for each of these stars (including temperature, surface gravity, metallicity and 15 chemical abundances) are estimated by the APOGEE Stellar Parameter
and Chemical Abundances Pipeline (ASPCAP; \citealt{Perez:2016})

\subsection{Training Samples}
\begin{enumerate}
\item[-] Describe the spectral-type training data \citep{West:2011}. One paragraph plus citations. Include any cuts we apply. \color{red}{Make diagnostic plots.}\color{black}

\item[-] Describe the Teff, Fe/H training data \citep{Mann:2015}. Same; Include any cuts. \color{red}{Make diagnostic plots.}\color{black}
\end{enumerate}

@@

\subsection{The Cannon: A Data-Driven Approach}
\begin{enumerate}
\item[-] A few-paragraph summary of what The Cannon does, with references (\citealt{Ness:2015}, \citealt{Casey:2016}, \citealt{Ho:2017a}, \citealt{Ho:2017b})
\end{enumerate}

>>>>

\usepackage{color}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{graphicx}

@@

The APOGEE survey \citep{Majewski:2015} of the Sloan Digital Sky Survey IV (SDSS-IV) is a high resolution (R$\sim$22,500), H-band (1.5-1.7$\mu$m) survey which has observed over 260,000 stellar spectra since its fourteenth data release (DR14; \citealt{Abolfathi:2017}). Fundamental parameters for each of these stars are estimated by the APOGEE Stellar Parameter and Chemical Abundances Pipeline (ASPCAP; \citealt{Perez:2016}), which employs a $\chi^2$ fitting procedure using the FERRE code to fit ATLAS9 radiative transfer models \citep{Castelli:2004} to determine atmospheric parameters, 15 chemical abundances and micro-turbulence parameters (T$_{\rm eff}$, log g, [Fe/H], [α/M], [C/M], [N/M]; \citealt{Meszaros:2012}).

APOGEE is primarily targeted for bright stellar populations, with de-reddened photometry and color cutoffs of $7 \leq H \leq 13.8$ and $[J-K]_0 \geq 0.5$ \citep{Zasowski:2013} with the objective of studying galactic composition and evolution. However numerous cool, main sequence sources have also been observed either accidentally, or as targets proposed by the M dwarf ancilliary survey \citep{Desphande:2013} (specific estimate based on Gaia?).

\subsection{Training Samples}

The Cannon is a \emph{generative model} which parameterizes the flux at each pixel of a spectrum in terms of a set of stellar labels (a flexible number of parameters chosen by the user; described in more detail in Section \ref{sec:cannon}). The model in this sense is used to \emph{transfer} labels from spectra which we know parameters for to those which we do not. Hence this data-driven approach effectively removes the challenges of physically modeling the atmosphere of a star (and common issues associated such as incomplete line lists or opacities), provided that we have a subset of spectra in the dataset with known (ideally very accurately measured) \emph{reference labels} that can be measured from other surveys/methods that are easier to infer from.

For the purpose of this study we have constructed two different training samples: first a one-dimensional \emph{spectral type model}, and second a two-dimensional \emph{physical parameter model} which describes the temperature and metallicity. The choice of training labels, dimensionality of our data set, and requirements for a good training set we discuss further in Section \ref{sec:discussion}.

The spectral type sample consists of 51 sources, spanning M0-M9 (Figure \ref{fig:train_dist}) cross-matched from the \citealt{West:2011} catalog of 78,841 M dwarfs from SDSS. For each source in the catalog, spectral types were determined both through an automated routine using The Hammer \citep{Covey:2007} and by visual inspection.

The physical parameter sample consists of 30 sources distributed over $3072 < T_{\rm eff} < 4131K$, and -0.48 $<$ [Fe/H] $<$ 0.49 (Figure \ref{fig:train_dist}) drawn from \citealt{Mann:2015}.
\color{red}{why logg not included as atmospheric parameter?}\color{black}

\begin{figure}[ht]
\plottwo{figures/west_train_dist.png}{figures/mann_train_dist.png}
\caption{Spectral type model: training label vs. test label.} \label{fig:west_selftest} \label{fig:train_dist}
\end{figure}

\subsection{Data Processing}
Describe ASPCAP spectra files and Ness continuum normalization procedure.

@@

\subsection{The Cannon: A Data-Driven Approach}

\begin{enumerate}
\item[-] A few-paragraph summary of what The Cannon does, with references (\citealt{Ness:2015}, \citealt{Casey:2016}, \citealt{Ho:2017a}, \citealt{Ho:2017b})
\end{enumerate}

Constructing such a model requires two steps: first the \emph{training step} in which the generative model is constructed at each pixel from the set of reference labels and fluxes, and second the \emph{test step} in which the model is applied to determine the labels of a spectrum

% The key advantages of The Cannon being:
% Data-driven models take away the challenge of directly infering labels from a survey, instead we \emph{transfer labels} from another survery that is more accurate or easier to model.
